Virtualisierung eines Echtzeit-Betriebssystems zur Steuerung eines Roboters mit Schwerpunkt auf die Einhaltung der Echtzeit

### 2024-08-15 Masterarbeit_Robotik_Pamuk.tex
Related Work 1.3 Papers read+

%
% FH Technikum Wien
% !TEX encoding = UTF-8 Unicode
%
% Erstellung von Master- und Bachelorarbeiten an der FH Technikum Wien mit Hilfe von LaTeX und der Klasse TWBOOK
%
% Um ein eigenes Dokument zu erstellen, müssen Sie folgendes ergänzen:
% 1) Mit \documentclass[..] einstellen: Master- oder Bachelorarbeit, Studiengang und Sprache
% 2) Mit \newcommand{\FHTWCitationType}.. Zitierstandard festlegen (wird in der Regel vom Studiengang vorgegeben - bitte erfragen)
% 3) Deckblatt, Kurzfassung, etc. ausfüllen
% 4) und die Arbeit schreiben (die verwendeten Literaturquellen in Literature.bib eintragen)
%
% Getestet mit TeXstudio mit Zeichenkodierung ISO-8859-1 (=ansinew/latin1) und MikTex unter Windows
% Zu beachten ist, dass die Kodierung der Datei mit der Kodierung des paketes inputenc zusammen passt!
% Die Kodierung der Datei twbook.cls MUSS ANSI betragen!
% Bei der Verwendung von UTF8 muss dnicht nur die Kodierung des Dokuments auf UTF8 gestellt sein, sondern auch die des BibTex-Files!
%
% Bugreports und Feedback bitte per E-Mail an latex@technikum-wien.at
%
% Versionen
% *) V0.7: 9.1.2015, RO: Modeline angepasst und verschoben
% *) V0.6: 10.10.2014, RO: Weitere Anpassung an die UK
% *) V0.5: 8.8.2014, WK: Literaturquellen überarbeitet und angepasst
% *) V0.4: 4.8.2014, WK: Initalversion in SVN eingespielt
%
\documentclass[MMR,Master,english]{twbook}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
%
% Hier biblatex & Biber konfigurieren; Vergessen Sie nicht, dass Sie biber verwenden müssen um eine Bibliothek zu erzeugen
%
\usepackage[backend=biber, style=numeric, sorting=none]{biblatex}
\addbibresource{Literature.bib}
%
% Bei Bedarf bitte hier die Syntax-Highlightings anpassen
%
% own 
\usepackage{hyperref}
\usepackage{float}
\usepackage{comment}
\usepackage{csquotes}
\usepackage{pgfplots}
\usepackage{soul}
\usepackage{color}
\usepackage{xcolor}  % Required for custom colors
\sethlcolor{blue}
\setcounter{secnumdepth}{3} % in text 
\setcounter{tocdepth}{3} % contents
\usepackage{longtable}
\usepackage{pdflscape}
\usepackage{rotating}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows}
\usetikzlibrary{positioning}
\usepackage[none]{hyphenat}

@@

\section{Real-Time Operating Systems}
\noindent The RTOS structure is visualized in Figure~\ref{fig:rtos_structure}.The core component of an RTOS that enables real-time capabilities is the kernel. It is responsible for managing system resources, scheduling tasks, and ensuring deterministic behavior. It employs preemptive scheduling mechanisms to allow high-priority tasks to preempt lower-priority tasks, ensuring that time-critical tasks are not delayed. Additionally, RTOS kernels are designed to allocate and manage memory resources efficiently and minimize interrupt latency, which is crucial for real-time applications that require immediate response to external events~\cite{wangRealtimeEmbeddedSystems2017}.

@@

\section{Application Context}\label{sec:application_context}

This section briefly defines the context and scope within which this master's thesis was composed, highlighting the specific contributions and requirements set by SIGMATEK GmbH \& Co KG~\cite{pixelartSIGMATEKKompletteAutomatisierungssysteme}.

@@

\section{Related Work and State of the Art}

\begin{comment}
Yocto primarily uses

At the beginning of the work, a detailed analysis of the individual virtualization options of Salamander 4 was carried out.

In particular, the virtualization performance of Ubuntu 22.04, Windows 10 and WSL under QEMU was compared.
\end{comment}

This master’s thesis builds upon a number of previous studies in the field of virtualization of real-time systems and the inherent latency. In this section, the most influential studies are presented and their relevance to the work is addressed.

\bigskip \noindent \citeauthor{perneelRealtimeCapabilitiesStandard2015} \cite{perneelRealtimeCapabilitiesStandard2015} mention in their work that Linux was initially developed as a general purpose operating system without real-time applications in mind. 
They talk about the evolution of real-time behavior in the standard Linux kernel and state that it has recently gained popularity among the real-time community, largely due to its open-source nature and stability. The authors explain how to transform Linux into a real-time operating system by implementing the PREEMPT-RT patches. The paper also emphasizes some kernel configurations that ensure that the system's real-time behavior is preserved during runtime. Even though these modifications to the kernel enable soft real-time performance, the authors underline the fundamental rule in real-time software, that latency improvements have a negative impact on throughput and kernel performance. This is due to the added overhead of the \texttt{CONFIG\_PREEMPT\_RT} option. They concluded with their Linux build and a testing application that Linux can be a viable option for real-time applications, if it is correctly configured.

@@

\bigskip \noindent Virtualization offers the great benefit of reducing overall hardware costs, since all non-critical activities and real-time control tasks can run on the same hardware. \citeauthor{scordinoRealTimeVirtualizationIndustrial2020} \cite{scordinoRealTimeVirtualizationIndustrial2020} take advantage of this, but also address the challenge of interference on shared hardware resources, which can degrade real-time performance.

\bigskip \noindent \citeauthor{maPerformanceTuningKVMbased} \cite{maPerformanceTuningKVMbased} \cite{junzhangPerformanceAnalysisKVMBased2010} describe in their work how KVM impacts the real-time performance, which has helped this work gain insight into the base overhead of virtualization. When a physical interrupt occurs while the guest RTOS is running, there are at least six stages that need to be traversed before the interrupt can be delivered to the guest RTOS. Once the interrupt is delivered to the guest RTOS, it incurs additional latency, called Interrupt Routine Time (IRT), since there is also the time between when an interrupt is recognized and the first instruction of the corresponding interrupt service routine (ISR) of the guest OS is started. These stages are the primary sources of latencies caused by KVM, they are breifly described in Figure \ref{fig:kvm_stages}.

\begin{figure}[H]
	\centering
	\begin{tikzpicture}[node distance = 0.45cm, auto,
		block/.style={
			rectangle,
			draw,
			fill=blue!20, 
			text width=35em, 
			rounded corners, 
			minimum height=3.5em,
			align=center,
			inner xsep=2em,
			inner ysep=0.5em 
		},
		line/.style={draw, -latex'}
	]
		% Place nodes
		\node [block] (init) {\textbf{Guest Code}\\ The guest OS is executing its normal operations.};
		\node [block, below=of init] (vmexit) {\textbf{VM Exit}\\ External interrupt triggers VM exit. Hardware stores interrupt info in VMCS.};
		\node [block, below=of vmexit] (aftervmexit) {\textbf{After VM Exit}\\ Control returns to kernel space. KVM retrieves interrupt info from VMCS,\\ marks it as pending for guest.};
		\node [block, below=of aftervmexit] (hostisr) {\textbf{Host ISR}\\ Host Linux handles the interrupt.};
		\node [block, below=of hostisr] (hostscheduling) {\textbf{Host Scheduling}\\ Host kernel invokes scheduler. Ideally, guest RTOS process isn't preempted.};
		\node [block, below=of hostscheduling] (beforevmentry) {\textbf{Before VM Entry}\\ KVM's kernel module prepares for VM entry. Detects pending interrupt,\\ injects virtual interrupt into guest by writing to VMCS.};
		\node [block, below=of beforevmentry] (vmentry) {\textbf{VM Entry}\\ Hardware emulates conventional interrupt handling for guest based on VMCS info.};
		\node [block, below=of vmentry] (guestirt) {\textbf{Guest IRT}\\ Guest's interrupt handler is invoked to service the interrupt.};
		% Draw edges
		\path [line] (init) -- (vmexit);
		\path [line] (vmexit) -- (aftervmexit);
		\path [line] (aftervmexit) -- (hostisr);
		\path [line] (hostisr) -- (hostscheduling);
		\path [line] (hostscheduling) -- (beforevmentry);
		\path [line] (beforevmentry) -- (vmentry);
		\path [line] (vmentry) -- (guestirt);
	\end{tikzpicture}
	\caption{Flowchart of operations in a virtualized environment}
	\label{fig:kvm_stages}
\end{figure}

@@

@@

\bigskip \noindent \citeauthor{deoliveiraDemystifyingRealTimeLinux} \cite{deoliveiraDemystifyingRealTimeLinux} critique the traditional \texttt{cyclictest} tool used for measuring scheduling latency, highlighting its limitations due to its black-box approach and lack of theoretical grounding. This is the main reason why this tool was not used to measure the scheduling latency in this work. Instead, the \texttt{latency} tool was used,as was determined in Chapter \ref{sec:application_context} .

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bigskip \noindent \textcolor{red}{Shielded Processors: Guaranteeing Sub-millisecond Response in Standard Linux by Steve Brosky and Steve Rotolo of Concurrent Computer Corporation explores the implementation and benefits of shielded processors in RedHawk Linux for real-time applications. The paper highlights the limitations of standard Linux in providing deterministic performance due to networking and graphics activities. Shielded CPUs, a model developed over a decade of experience, offload operating system overhead to a subset of CPUs, ensuring more deterministic execution and interrupt response times. The authors describe the modifications made to the Linux kernel, including CPU affinity settings and the disabling of local timer interrupts, to create a shielded environment. Benchmark tests demonstrate significant improvements in execution determinism and interrupt response times, with jitter reduced to 1.87\% on shielded CPUs compared to 26.17\% on standard Linux. The paper concludes that shielded processors can guarantee interrupt response times of less than 30 microseconds, even under heavy system load, making them suitable for applications with hard real-time deadlines. The work underscores the potential of using commodity operating systems like Linux for high-performance real-time applications by leveraging shielded processors.}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

%
% FH Technikum Wien
% !TEX encoding = UTF-8 Unicode
%
% Erstellung von Master- und Bachelorarbeiten an der FH Technikum Wien mit Hilfe von LaTeX und der Klasse TWBOOK
%
% Um ein eigenes Dokument zu erstellen, müssen Sie folgendes ergänzen:
% 1) Mit \documentclass[..] einstellen: Master- oder Bachelorarbeit, Studiengang und Sprache
% 2) Mit \newcommand{\FHTWCitationType}.. Zitierstandard festlegen (wird in der Regel vom Studiengang vorgegeben - bitte erfragen)
% 3) Deckblatt, Kurzfassung, etc. ausfüllen
% 4) und die Arbeit schreiben (die verwendeten Literaturquellen in Literature.bib eintragen)
%
% Getestet mit TeXstudio mit Zeichenkodierung ISO-8859-1 (=ansinew/latin1) und MikTex unter Windows
% Zu beachten ist, dass die Kodierung der Datei mit der Kodierung des paketes inputenc zusammen passt!
% Die Kodierung der Datei twbook.cls MUSS ANSI betragen!
% Bei der Verwendung von UTF8 muss dnicht nur die Kodierung des Dokuments auf UTF8 gestellt sein, sondern auch die des BibTex-Files!
%
% Bugreports und Feedback bitte per E-Mail an latex@technikum-wien.at
%
% Versionen
% *) V0.7: 9.1.2015, RO: Modeline angepasst und verschoben
% *) V0.6: 10.10.2014, RO: Weitere Anpassung an die UK
% *) V0.5: 8.8.2014, WK: Literaturquellen überarbeitet und angepasst
% *) V0.4: 4.8.2014, WK: Initalversion in SVN eingespielt
%
\documentclass[MMR,Master,english]{twbook}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
%
% Hier biblatex & Biber konfigurieren; Vergessen Sie nicht, dass Sie biber verwenden müssen um eine Bibliothek zu erzeugen
%
\usepackage[backend=biber, style=numeric, sorting=none]{biblatex}
\addbibresource{Literature.bib}
%
% Bei Bedarf bitte hier die Syntax-Highlightings anpassen
%
% own 
\usepackage{hyperref}
\usepackage{float}
\usepackage{comment}
\usepackage{csquotes}
\usepackage{pgfplots}
\usepackage{soul}
\usepackage{color}
\usepackage{xcolor}  % Required for custom colors
\sethlcolor{blue}
\setcounter{secnumdepth}{3} % in text 
\setcounter{tocdepth}{3} % contents
\usepackage{longtable}
\usepackage{pdflscape}
\usepackage{rotating}
\usepackage{tikz}
\usepackage{setspace}
\usetikzlibrary{shapes,arrows}
\usetikzlibrary{positioning}
\usepackage[none]{hyphenat}

@@

\section{Real-Time Operating Systems}
\noindent The RTOS structure is visualized in Figure~\ref{fig:rtos_structure}.The core component of an RTOS that enables real-time capabilities is the kernel. It is responsible for managing system resources, scheduling tasks, and ensuring deterministic behavior \cite{malallahComprehensiveStudyKernel2021}. It employs preemptive scheduling mechanisms to allow high-priority tasks to preempt lower-priority tasks, ensuring that time-critical tasks are not delayed. Additionally, RTOS kernels are designed to allocate and manage memory resources efficiently and minimize interrupt latency, which is crucial for real-time applications that require immediate response to external events~\cite{wangRealtimeEmbeddedSystems2017}.

@@

\section{Application Context}\label{sec:application_context}
This section briefly defines the context and scope within which this master's thesis was composed, highlighting the specific contributions and requirements set by SIGMATEK GmbH \& Co KG~\cite{pixelartSIGMATEKKompletteAutomatisierungssysteme}.

@@

\section{Related Work and State of the Art}
This master’s thesis builds upon a number of previous studies in the field of virtualization of real-time systems and the inherent latency. In this section, the most influential studies are presented and their relevance to the work is addressed.

\bigskip \noindent \citeauthor{perneelRealtimeCapabilitiesStandard2015} \cite{perneelRealtimeCapabilitiesStandard2015} mention in their work that Linux was initially developed as a general purpose operating system without real-time applications in mind. They talk about the evolution of real-time behavior in the standard Linux kernel and state that it has recently gained popularity among the real-time community, largely due to its open-source nature and stability. The authors explain how to transform Linux into a real-time operating system by implementing the PREEMPT-RT patches. The paper also emphasizes some kernel configurations that ensure that the system's real-time behavior is preserved during runtime. Even though these modifications to the kernel enable soft real-time performance, the authors underline the fundamental rule in real-time software, that latency improvements have a negative impact on throughput and kernel performance. This is due to the added overhead of the \texttt{CONFIG\_PREEMPT\_RT} option. They concluded with their Linux build and a testing application that Linux can be a viable option for real-time applications, if it is correctly configured. Similarly, \citeauthor{adamRealTimePerformanceResponse2021} \cite{adamRealTimePerformanceResponse2021} applied the PREEMPT\_RT patch on both Raspberry Pi 3 and BeagleBone Black and achieved significantly lower latencies compared to standard Linux kernels.

@@

\bigskip \noindent \citeauthor{taccariEmbeddedRealTimeVirtualization2014} \cite{taccariEmbeddedRealTimeVirtualization2014}, \citeauthor{sandstromVirtualizationTechnologiesEmbedded2013} \cite{sandstromVirtualizationTechnologiesEmbedded2013}, and \citeauthor{guStateoftheArtSurveyRealTime2012} \cite{guStateoftheArtSurveyRealTime2012} all discuss and compare the current state-of-the-art in virtualization technologies on embedded real-time platforms.

highlighting similar advancements and challenges in their respective works.

\bigskip \noindent 
\citeauthor{garcia-vallsChallengesRealtimeVirtualization2014} \cite{garcia-vallsChallengesRealtimeVirtualization2014}

\citeauthor{cinqueVirtualizingMixedCriticalitySystems2022} \cite{cinqueVirtualizingMixedCriticalitySystems2022}

\citeauthor{reghenzaniRealTimeLinuxKernel2020} \cite{reghenzaniRealTimeLinuxKernel2020}

\bigskip \noindent Virtualization offers the great benefit of reducing overall hardware costs, since all non-critical activities and real-time control tasks can run on the same hardware. \citeauthor{scordinoRealTimeVirtualizationIndustrial2020} \cite{scordinoRealTimeVirtualizationIndustrial2020} take advantage of this, but also address the challenge of interference on shared hardware resources, which can degrade real-time performance. In addition, \citeauthor{javierperezHowRealTime2022} \cite{javierperezHowRealTime2022} state that hardware-based PLCs are costly to maintain and lack the flexibility needed for modern, resource-intensive applications like Machine Learning (ML) or Artificial Intelligence (AI).

\bigskip \noindent \citeauthor{maPerformanceTuningKVMbased} \cite{maPerformanceTuningKVMbased} \cite{junzhangPerformanceAnalysisKVMBased2010} describe in their work how KVM impacts the real-time performance, which has helped this work gain insight into the base overhead of virtualization. When a physical interrupt occurs while the guest RTOS is running, there are at least six stages that need to be traversed before the interrupt can be delivered to the guest RTOS. Once the interrupt is delivered to the guest RTOS, it incurs additional latency, called Interrupt Routine Time (IRT), since there is also the time between when an interrupt is recognized and the first instruction of the corresponding interrupt service routine (ISR) of the guest OS is started. These stages are the primary sources of latencies caused by KVM, they are briefly described in Figure \ref{fig:kvm_stages}.

\begin{figure}[H]
	\centering
	\begin{tikzpicture}[node distance = 0.35cm, auto, % Adjusted node distance
		block/.style={
			rectangle,
			draw,
			fill=blue!20, 
			text width=35em, 
			rounded corners, 
			minimum height=3.5em,
			align=center,
			inner sep=0.2em % Adjusted inner separation
		},
		line/.style={draw, -latex'}
	]
		% Place nodes
		\node [block] (init) {\scriptsize \textbf{Guest Code}\\  The guest OS is executing its normal operations.};
		\node [block, below=of init] (vmexit) {\scriptsize \textbf{VM Exit}\\  External interrupt triggers VM exit. Hardware stores interrupt info in VMCS.};
		\node [block, below=of vmexit] (aftervmexit) {\scriptsize \textbf{After VM Exit}\\  Control returns to kernel space. KVM retrieves interrupt info from VMCS,\\\vspace{-0.5em} marks it as pending for guest.};
		\node [block, below=of aftervmexit] (hostisr) {\scriptsize \textbf{Host ISR}\\  Host Linux handles the interrupt.};
		\node [block, below=of hostisr] (hostscheduling) {\scriptsize \textbf{Host Scheduling}\\  Host kernel invokes scheduler. Ideally, guest RTOS process isn't preempted.};
		\node [block, below=of hostscheduling] (beforevmentry) {\scriptsize \textbf{Before VM Entry}\\  KVM's kernel module prepares for VM entry. Detects pending interrupt,\\\vspace{-0.5em} injects virtual interrupt into guest by writing to VMCS.};
		\node [block, below=of beforevmentry] (vmentry) {\scriptsize \textbf{VM Entry}\\  Hardware emulates conventional interrupt handling for guest based on VMCS info.};
		\node [block, below=of vmentry] (guestirt) {\scriptsize \textbf{Guest IRT}\\  Guest's interrupt handler is invoked to service the interrupt.};
		% Draw edges
		\path [line] (init) -- (vmexit);
		\path [line] (vmexit) -- (aftervmexit);
		\path [line] (aftervmexit) -- (hostisr);
		\path [line] (hostisr) -- (hostscheduling);
		\path [line] (hostscheduling) -- (beforevmentry);
		\path [line] (beforevmentry) -- (vmentry);
		\path [line] (vmentry) -- (guestirt);
	\end{tikzpicture}
	\caption{Flowchart of operations in a virtualized environment}
	\label{fig:kvm_stages}
\end{figure}

@@

\bigskip \noindent \citeauthor{broskyShieldedProcessorsGuaranteeing2003} \cite{broskyShieldedProcessorsGuaranteeing2003} report in their work about the shielded CPU model for enhancing real-time performance in symmetric multiprocessor systems. By dedicating one or more CPUs to high-priority processes and interrupts, the goal is to provide deterministic execution of real-time applications and interrupt responses. The authors mention that there is no user interface for setting process CPU affinity in the standard Linux, while there is one for setting interrupt CPU affinity via the \texttt{/proc/irq/*/smp\_affinity} files. These files were used in this masther's thesis. According to \citeauthor{broskyShieldedProcessorsGuaranteeing2003}, the shielded CPU model is especially useful for tasks that require guaranteed interrupt response times, very fast interrupt responses and high-frequency, deterministic execution.

@@

\bigskip \noindent \citeauthor{deoliveiraDemystifyingRealTimeLinux} \cite{deoliveiraDemystifyingRealTimeLinux} critique the traditional \texttt{cyclictest} tool used for measuring scheduling latency, highlighting its limitations due to its black-box approach and lack of theoretical grounding. This is the main reason why this tool was not used to measure the scheduling latency in this work. Instead, the \texttt{latency} tool was used,as was determined in Chapter \ref{sec:application_context}.

@@
